HO unification from object language to meta language

### 2024-03-25 paper/main.tex
more5

\begin{elpicode}
decision (all A x\ Pm x) :- link Pm P A, finite A,   ~\customlabel{clause:r3a}{(r3a)}~
  pi x\ decision (app[P, x]).
\end{elpicode}

@@

\noindent
After unifying the head of rule \ref{clause:r3a} with the goal, Elpi runs
the premise <<\elpiIn{link Pm A P}>> that is in charge of bringing the
assignment for \elpiIn{Pm} back to the domain \elpiIn{tm} of Coq terms:

@@

\begin{invariant}[Link left hand side] The left hand side of a new link
  is a variable.
\end{invariant}

@@

XXX links are update unlike section 2
\input{code/hstep}

TODOD: discuss signature of progress

@@

\noindent
Note thar \elpiIn{((A ~\Ue~B) C D)} is syntactic sugar for
\elpiIn{((~\Ue~) A B C D)}.

TODO: why it terminates

@@

\input{code/decomp}
\input{code/decompm}

>>>>

\begin{elpicode}
decision (all A x\ Pm x) :- decomp Pm P A, finite A,   ~\customlabel{clause:r3a}{(r3a)}~
  pi x\ decision (app[P, x]).
\end{elpicode}

@@

\noindent
After unifying the head of rule \ref{clause:r3a} with the goal, Elpi runs
the premise <<\elpiIn{decomp Pm A P}>> that is in charge of bringing the
assignment for \elpiIn{Pm} back to the domain \elpiIn{tm} of Coq terms:

@@

\begin{invariant}[Link left hand side]\ref{inv:linklhs}
  The left hand side of a new link
  is a variable.
\end{invariant}

@@

A step in \Ho consists in unifying two terms and reconsidering all
links for progress. If any of the two tasks fail we say that the entire step
fails, and it is at this granularity that we can relate steps in the
two languages.

\input{code/hstep}

\noindent
Note tha the notation \elpiIn{((A ~\Ue~B) C D)} is syntactic sugar for
\elpiIn{((~\Ue~) A B C D)}.

Reconsidering links is a fixpoint, since the progress of a link can update the
substitution and in turn enable another link to progress.

@@

\noindent
In the base compilation scheme \elpiIn{progress1} is the identity
on both the links and the substitution, so the fixpoint trivially terminates.
Sections \ref{sec:eta} and \ref{sec:beta} add rules to \elpiIn{progress1}
and justify why the don't hinder termination.

TODO: discuss occur check

@@

Decompiling the substitution requires to first force the
progress of links and then allocating new unassigned variables
in the substitution for \Fo{} and finally decompiling all
assignments. Note that \ref{inv:linklhs} and the
occur check allows us to update the subst.

\input{code/decompile}

\noindent
Decompiling an assignment requires to turn abstractions into
lambdas. For aestetic purposes we also eta-contract the result
(not needed since Fo equality can do that)

\input{code/decompm}

\noindent
Finally decompiling a term is trivial, now that we have an extended
mapping containing all unassigned variables \Ue may have introduced.

\input{code/decomp}

\noindent
Note that we use beta to build fapp nodes when needed (if Ag is empty
no \elpiIn{fapp} node should appear).
